---
title: "Download GDSC, CTRP, and TCGA Data"
output:
  html_document:
    df_print: paged
author:
  - "Robert F. Gruener"
date: "date last edited: `r format(Sys.time(), '%d %B, %Y')`"
---

## Objectives:

The goal of this file is to: download and install the necessary packages and files into the R project. Specifically: 

 * Install pRRophetic and it's dependencies
 * Download the cell line screening data
 * Download the RNA-Seq information
 * Download the TCGA dataset
 
Important note, certain other files are needed as well: "pRRophetic_0.5.tar.gz" will be needed to perform the imputations. "All Cell Line Infor.xlsx" , and  "Breast_CCL_Subtypes.csv". Are needed for the TNBC analysis, but those downloads are separate. 

pRRophetic_0.5.tar.gz can be downloaded [here](https://osf.io/dwzce/?action=download). Place this in the repository or change the location of the installation in the first code chunk. 

"All Cell Line Infor.xlsx" , and  "Breast_CCL_Subtypes.csv". Are needed for the TNBC analysis, but those downloads are separate.  

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, cache = T)
```


### Installing required packages (pRRophetic must be downloaded and put in wd or project home)

Again to install pRRophetic, it will need to be downloaded and placed in the repo manually/separately. 

```{r install and load pRRophetic and dependencies, results= "hide"}

#install BiocManager if needed
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
#install pRRophetic dependencies if needed
if (!requireNamespace(c("car", "ridge", "preprocessCore", "genefilter", "sva"), quietly = TRUE)) 
    BiocManager::install(c("car", "ridge", "preprocessCore", "genefilter", "sva"))

#install pRRophetic from tar.gz file (must be placed in repo seperately, or the location needs to be changed)
if (!requireNamespace("pRRophetic", quietly = TRUE)) 
    install.packages("pRRophetic_0.5.tar.gz", repos = NULL, type = "source")

#install other packages that will be used in this project
req_packages <- c("readxl", "TCGAbiolinks", "tidyverse", "magrittr", "ggpubr", "broom", "grid", "ggrepel", "lsr", "svglite")
needed_packs <- req_packages[!(req_packages %in% installed.packages())]
if(length(needed_packs)) BiocManager::install(needed_packs)

dir.create("./DataIn")

```

### Downloading Cell Line Data


#### Download the data from GDSC. 

Downloading the GDSC2 data from Sanger. Last attempted 5/2020, downloaded data from July 2019 update: Release 8.0. 


NB: Can find the GDSC expression data [at this link](https://www.cancerrxgene.org/gdsc1000/GDSC1000_WebResources//Data/preprocessed/Cell_line_RMA_proc_basalExp.txt.zip )
However, this information wasn't updated any since 2016 and the 2016 data is already packaged into this version of pRRophetic as cgp2016ExprRma, so we will just call the data from pRRophetic.(True as of 9/2019) 

```{r downloading GDSC IC50 data}

#downloading the GDSC2 8.0 release
dir.create("./DataIn/GDSC")
download.file(url = "ftp://ftp.sanger.ac.uk/pub/project/cancerrxgene/releases/release-8.0/GDSC2_fitted_dose_response_17Jul19.xlsx",
                destfile = "./DataIn/GDSC/GDSC_IC50_July2019.xlsx", mode = "wb")

```


Downloads the data from CTRP and CCLE. Data can also be accessed from these links:

CTRPv2 Data from: ftp://caftpd.nci.nih.gov/pub/OCG-DCC/CTD2/Broad/CTRPv2.0_2015_ctd2_ExpandedDataset/
CCLE Data from: https://portals.broadinstitute.org/ccle/data


```{r downloading CTRP and CCLE data}
dir.create("./DataIn/CTRPv2")
dir.create("./DataIn/CCLE")

#CTRP Data
temp.zip <- tempfile()
download.file(url = "ftp://caftpd.nci.nih.gov/pub/OCG-DCC/CTD2/Broad/CTRPv2.0_2015_ctd2_ExpandedDataset/CTRPv2.0_2015_ctd2_ExpandedDataset.zip",
              "./DataIn/CTRP.zip") 
unzip("./DataIn/CTRP.zip", exdir = "./DataIn/CTRPv2")
unlink("./DataIn/CTRP.zip")

#Download CCLE Data 
download.file(url = "https://data.broadinstitute.org/ccle/CCLE_DepMap_18Q2_RNAseq_RPKM_20180502.gct", 
              destfile = "./DataIn/CCLE/CCLE_RPKM.gct")


```

### Downloading the TCGA data from FireBrowse


```{r downloading TCGA data}
#this should be updated to the GDC instead of Firebrowse at some point
dir.create("./DataIn/TCGA")
download.file("http://gdac.broadinstitute.org/runs/stddata__2015_08_21/data/BRCA/20150821/gdac.broadinstitute.org_BRCA.Merge_rnaseqv2__illuminahiseq_rnaseqv2__unc_edu__Level_3__RSEM_genes__data.Level_3.2015082100.0.0.tar.gz", "./DataIn/TCGA/TCGA_BRCA_RNA-seq.tar.gz")
untar("./DataIn/TCGA/TCGA_BRCA_RNA-seq.tar.gz", exdir = "./DataIn/TCGA/")
#having issues with the untar file name being too long, renaming
file.rename(from = "./DataIn/TCGA/gdac.broadinstitute.org_BRCA.Merge_rnaseqv2__illuminahiseq_rnaseqv2__unc_edu__Level_3__RSEM_genes__data.Level_3.2015082100.0.0/", to = "./DataIn/TCGA/RNA-seq" )
unlink("./DataIn/TCGA/TCGA_BRCA_RNA-seq.tar.gz")


```


```{r}
print(sessionInfo())
```

