---
title: 'Imputations: Build models using the CCLE/CTRPv2 datasets to impute drug response in TCGA BRCA Tumors'
output: html_notebook 
author:
  - "Robert F. Gruener"
date: "date last edited: `r format(Sys.time(), '%d %B, %Y')`"
---
```{r, include=FALSE}
knitr::opts_chunk$set(warning = F, message = FALSE, cache = T)
```
### Objectives

Do the imputations

 * Set the parameters
 * Load the test expression data (TCGA Breast Cancer)
 * Load the CTRP/CCLE Training expression and drug data
 * Generate the models and impute using the pRRophetic framework 

Important: Set the parameters appropriately, then everything else *should* run smoothly if you run the enitre file. 



### First, load packages and set parameters

The necessary packages and files should have been download in file 01_Downlaod

```{r loading packages and setting parameters, results= "hide"}
set.seed(12345)
#Loading Packages
needed_packages <- c("tidyverse", "readxl", "glmnet", "gdata", "illuminaHumanv4.db" ,"car", "ridge", "preprocessCore", "genefilter", "sva", "pRRophetic")
lapply(needed_packages, require, character.only = TRUE)

###################### Set Parameters ##########################
# Set location and filename  for list/matrix R.data output (ctrp_list_output_name, ctrp_mat_output_name)

dir.create("DataOut")
#matrix_output_name <- "./DataOut/2015CTRP-Models_on_TCGA_BRCA.RData"

# Finally, set parameters of pRRophetic's calcphenotype function 
#batchCorrect options: "eb" for ComBat, "qn" for quantiles normalization, "standardize", or "none"
batchCorrect <- "eb"
#powerTransformPhenotype: does a power transformation to the phenotype data, recommended over log transformations by the literture 
powerTransformPhenotype <- TRUE
#removeLowVaryingGenes: percentage of low varying genes to remove, default is at 0.2 
removeLowVaryingGenes <- 0.2
#removeLowVaryingGenesFrom: where should it remove low varying genes from "homogenizeData" or "rawData"
removeLowVaringGenesFrom <- "homogenizeData"
#minNumSamples: minimum number of training samples, set to 30, but this shouldn't be an issue since the there are many samples 
minNumSamples = 30
#selection: how to deal with duplicated gene IDs, -1 for ask user, 1 for summarize by mean, 2 for disregard duplicates
selection <- 1
#printoutput: should it print outputs, default is TRUE
printOutput = TRUE


```

### Second, load the TCGA RNA-Seq data, remove normal samples

```{r Load RNA-Seq}
############ Load Test Expression data (as Matrix) ###################
tpmDatMat_tpm_logged <- read.delim("./DataIn/TCGA/TCGA_Breast_TPM.txt", sep = "\t", stringsAsFactors = F)

#remove adjacent normal RNA-seq samples
testExprData <- tpmDatMat_tpm_logged[,-grep(pattern = ".11A.|.11B.|.10A.|.10B.", colnames(tpmDatMat_tpm_logged))]

```

### Third, load and edit the CTRP/CCLE data:

Perform imputations using CTRP/CCLE datasets as the training dataset

```{r CTRP Imputations, message = FALSE}

#Loads training phenotype data 
CTRPv2_AUC <- read.delim("./DataIn/CTRPv2/CTRPv2_AUC_clean.txt", sep = "\t", stringsAsFactors = F)
#Altering cell line names that start with a number to start with an X to match rownames of expression data
start.with.number <- grep("^[[:digit:]]", CTRPv2_AUC$CCL_Name)
CTRPv2_AUC$CCL_Name[start.with.number] <- paste("X", CTRPv2_AUC$CCL_Name[start.with.number], sep = "")
trainingPhenData <- CTRPv2_AUC
possibleDrugs <- unique(trainingPhenData$cpd_name)

#Load Training Expression data, convert it to a matrix
CTRPv2_RNAseq_TPM <- read.delim("./DataIn/CTRPv2/CTRPv2_RNAseq_TPM_clean.txt",  sep = "\t", stringsAsFactors = F)
convert.to.matrix <- function(df){
  mat <- as.matrix(df[,-1])
  rownames(mat) <- c(t(df[,1]))
  return(mat)
  }
ExprData <- convert.to.matrix(CTRPv2_RNAseq_TPM)



```


### Do the actual imputations 

```{r, autodep= TRUE}
#make a list of all possible drugs predictions
CTRPDrugPredictions <- list()

########make a for loop for each drug seperate#################
for(i in 1:length(possibleDrugs)) { 
  drug <- possibleDrugs[i]
  temp <- trainingPhenData[trainingPhenData$cpd_name == drug,] ## Makes temp matrix to extract the AUCs easier
  AUCs <- as.numeric(temp$Avg_AUC)
  names(AUCs) <- temp$CCL_Name
  commonCellLines <- colnames(ExprData)[colnames(ExprData) %in% names(AUCs)]
  AUCsOrd <- AUCs[commonCellLines]
  trainDataOrd <- ExprData[, commonCellLines]
  trainDataOrd <- trainDataOrd[(rowSums(trainDataOrd) != 0), ]

  CTRPDrugPredictions[[i]] <- calcPhenotype(trainingExprData =  trainDataOrd, 
                                trainingPtype = AUCsOrd, testExprData = testExprData, batchCorrect = batchCorrect, 
                                powerTransformPhenotype = powerTransformPhenotype, removeLowVaryingGenes = removeLowVaryingGenes, 
                                minNumSamples = minNumSamples, selection = selection, 
                                printOutput = printOutput, removeLowVaringGenesFrom = removeLowVaringGenesFrom)
} #close for loop

names(CTRPDrugPredictions) <- possibleDrugs
CTRPDrugPredictions_mat <- do.call(rbind, CTRPDrugPredictions)
save(CTRPDrugPredictions_mat, file= matrix_output_name)
```



```{r}
print(sessionInfo())
```




