---
title: "Analyzing TCGA (patient) Imputation Data"
output: html_notebook
---

To work on: 
   * need to (most likely) remove the eta.sq() results
   * need to redo the TNBC analysis, removing eta.sq() component
   * need to decide on clincal or PAM50 subtyping 


Analysis outline: 

  * Install TCGAbiolinks to get TCGA clinical data
  * Format BRCA predictions, merge with clinical data
  * Perform t-tests and correct for multiple testing 

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```


```{r install packages, include= FALSE}

if (!requireNamespace("TCGAbiolinks", quietly = TRUE))
    BiocManager::install("TCGAbiolinks")

if (!requireNamespace(c("broom", "grid", "ggrepel", "lsr"), quietly = TRUE))
    install.packages("broom", "grid", "ggrepel", "lsr")

library(TCGAbiolinks)
library(tidyverse)
library(broom)
library(readxl)
library(grid)
library(ggrepel)

theme_set(theme_bw())

```

```{r download TCGA clinical data}
query <- GDCquery(project = "TCGA-BRCA", 
                  data.category = "Clinical",
                  data.type = "Clinical Supplement", 
                  data.format = "BCR Biotab")
GDCdownload(query)
clinical.BCRtab.all <- GDCprepare(query)
names(clinical.BCRtab.all)
BRCA_clinical <- clinical.BCRtab.all$clinical_patient_brca %>%
    dplyr::select(bcr_patient_barcode, er_status_by_ihc, er_status_ihc_Percent_Positive, pr_status_by_ihc, 
                  pr_status_ihc_percent_positive, her2_status_by_ihc, her2_fish_status, ) %>%
    dplyr::slice(-1:-2) %>% 
    type_convert() %>% 
    mutate(HER2_Subtype = if_else(condition = her2_fish_status %in% c("Positive", "Negative"), her2_fish_status, her2_status_by_ihc)) %>% 
    filter(er_status_by_ihc %in% c("Positive", "Negative"), 
           pr_status_by_ihc %in% c("Positive", "Negative"), 
           HER2_Subtype %in% c("Positive", "Negative")) %>% 
    mutate(Clin_Subtype = if_else(condition = er_status_by_ihc == "Positive" | pr_status_by_ihc == "Positive", 
                                  true = if_else(condition = HER2_Subtype == "Positive", true = "ER/PR+, HER2+", false = "ER/PR+"),
                                  false = if_else(condition = HER2_Subtype == "Positive", true = "HER2+", false = "TNBC"))) %>% 
    mutate(TNBC_status = if_else(Clin_Subtype == "TNBC", "TNBC", "RPBC"))

TCGA_subs <- TCGAquery_subtype(tumor = "brca") #easily get TCGA PAM50 status this way

BRCA_pam50 <- TCGA_subs %>% select(patient, PAM50 = BRCA_Subtype_PAM50) %>% filter(PAM50 != "NA")

BRCA_clinical <- full_join(BRCA_clinical, BRCA_pam50, by = c("bcr_patient_barcode" = "patient")) 

```

This table shows the clinical breakdown by subtype. 

```{r}
table(BRCA_clinical$Clin_Subtype)
table(BRCA_clinical$PAM50)

```

Ki-67 staining is not available in TCGA, so we cannot do the traditional Lum A and Lum B groups for the clinical subtyping of ER+ tumors. However, this doesn't impact us too much as there we are focused on TNBC which can be easily defined in our dataset.  

### Identifying Compounds Predicted To Be More Effective in TNBC

First we need to load and format the prediction dataset, then we merge it with the clinical data we previously downloaded

```{r format predictions, include = FALSE}
load("./DataOut/2015CTRP-Models_on_TCGA_BRCA.RData")
BRCA_preds <- as_tibble(t(CTRPDrugPredictions_mat), rownames = "Patient_ID") 
BRCA_preds <- BRCA_preds %>% 
  filter(grepl(Patient_ID, pattern = "\\.01A\\.|\\.01B\\.")) %>% 
  mutate(Patient_ID = gsub(x = strtrim(Patient_ID, width = 12), pattern = "\\.", replacement = "-")) %>% 
  gather(key = "Drug", value = "PSS", -Patient_ID)

#join with clinical information
preds.subtype <- left_join(BRCA_preds, BRCA_clinical, by = c("Patient_ID" = "bcr_patient_barcode")) %>% 
  drop_na(TNBC_status, PAM50)

```

To look for compounds that are more effective in TNBC vs other breast cancers (i.e. receptor positive or RPBC), we can subset the patients into their corresponding TNBC status and perform t-tests on the imputed drug data for each drug. Then we can correct for multiple test corrections. We can then see how many drugs are expected to be more effective in TNBC compared to RPBCs, which is shown in the below table.  

```{r Statistical testing}
t.tst_preds.subtype <- preds.subtype %>% 
  group_by(Drug) %>% 
  do(tidy(t.test(PSS~TNBC_status, data = .))) %>%
  ungroup() %>% 
  mutate(bonf_p.value = p.adjust(p.value, method = "bonferroni")) %>% 
  mutate(log10_p.correct = -log10(bonf_p.value)) 

t.tst_preds.subtype %>% filter(bonf_p.value < 0.001) %>% mutate(Sensitive = if_else(estimate > 0, "TNBC", "RPBC")) %>% count(Sensitive)

#We could also look at the results via ANOVA and look for the most significantly different values between the clinical subgroups described before or the PAM50 subgroups. I'm keeping this analysis here but some of this may become supplemental.  
aov_preds.subtype <- preds.subtype %>%
  group_by(Drug) %>%
  do(tidy(aov(.$PSS ~ .$PAM50))) %>%
  ungroup() %>% 
  filter(term == ".$PAM50") %>%
  mutate(bonf_p.value = p.adjust(p.value, method = "bonferroni"))

clin_aov_preds.subtype <- preds.subtype %>%
  group_by(Drug) %>%
  do(tidy(aov(.$PSS ~ .$Clin_Subtype))) %>%
  ungroup() %>% 
  filter(term == ".$Clin_Subtype") %>%
  mutate(bonf_p.value = p.adjust(p.value, method = "bonferroni"))



```


As a note: 
Based on t-tests, even with a bonferroni correction, most of the values are statistically signifcant. From this table `r sum(t.tst_preds.subtype$bonf_p.value < 0.001/length(unique(t.tst_preds.subtype$Drug)), na.rm = TRUE)`% are statistically significant at a corrected p-value of 0.001 out of the `r length(unique(t.tst_preds.subtype$Drug))`total drugs. This does indicte that we are getting an enrichment of likely false positive results, so the best thing to do would be to just consider the most significant results, essentially placing a much higher significant cutoff threshold. This should be investigated in future analysis. 

```{r, echo= FALSE}

t.tst_preds.subtype %>% 
  ggplot() +
    geom_histogram(mapping = aes(x = log10_p.correct)) +
    geom_vline(xintercept = -log10(0.001), color = "red") +
    labs(title = "Histogram of T-test Results by Signifcance", 
         x = "Number of Drugs", 
         y = "-log10(bonferroni-adjusted p-value")

```

Now we can create a volcano plot with the t-test results and look for the most significant results: 

```{r volcano plot of Estimate vs p-value}

t.tst_preds.subtype %>% 
  arrange(desc(log10_p.correct)) %>% 
  dplyr::slice(1:5) %>% 
  ggplot(mapping = aes(x = estimate, y = log10_p.correct, label = Drug)) + 
    geom_point(data = t.tst_preds.subtype) +
    geom_point(color = "red", size = 2.2) + 
    geom_text_repel(nudge_y = 3) + 
    labs(title = "Volcano Plot of TNBC vs RPBC T-Test Results \n P-values and Estimate Scores for All Drugs", 
         y = "-log10(p.value)", 
         x = "Mean Difference (Imputed Effect RPBC - TNBC)", 
         caption = "Drugs More Effective in RPBC                                                                                                            Drugs More Effective in TNBC") + 
    theme(plot.title = element_text(hjust = 0.5))
    
```


### Postive controls and 


Let us see if this recapitulates certain expected subtype specific response. 


```{r graphing function}

PAM50_ISS_grapher <- function(drug){
  grob_annot <- grobTree(textGrob(paste("p-value =", 
                                      format(aov_preds.subtype$bonf_p.value[aov_preds.subtype$Drug == drug], 
                                             trim = T, digits = 4)), 
                                x = 0.73, y = 0.95, hjust = 0, gp = gpar(fontface = "italic")))
  print(preds.subtype %>% 
    filter(Drug == drug) %>% 
    ggplot(aes(x = PAM50, y = PSS, fill = PAM50)) + 
      geom_boxplot() +
      labs(title = paste(drug, "'s Imputed Sensitivity in TCGA BRCA Cohort by PAM50 Subtype", sep = ""),
           y = "Imputed Sensitivity Score") +
      theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5))+
      annotation_custom(grob_annot))
      
}

Clin.sub_ISS_grapher <- function(drug){
  grob_annot <- grobTree(textGrob(paste("p-value =", 
                                      format(clin_aov_preds.subtype$bonf_p.value[clin_aov_preds.subtype$Drug == drug], 
                                             trim = T, digits = 4)), 
                                x = 0.73, y = 0.95, hjust = 0, gp = gpar(fontface = "italic")))
  print(preds.subtype %>% 
    filter(Drug == drug) %>% 
    ggplot(aes(x = Clin_Subtype, y = PSS, fill = Clin_Subtype)) + 
      geom_boxplot() +
      labs(title = paste(drug, "'s Imputed Sensitivity in TCGA BRCA Cohort by Clinical Subtype", sep = ""),
           y = "Imputed Sensitivity Score") +
      theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5))+
      annotation_custom(grob_annot))
      
}

TNBC_ISS_grapher <- function(drug){
  grob_annot <- grobTree(textGrob(paste("p-value =", 
                                      format(t.tst_preds.subtype$bonf_p.value[t.tst_preds.subtype$Drug == drug], 
                                             trim = T, digits = 4)), 
                                x = 0.73, y = 0.95, hjust = 0, gp = gpar(fontface = "italic")))
  print(preds.subtype %>% 
    filter(Drug == drug) %>% 
    ggplot(aes(x = TNBC_status, y = PSS, fill = TNBC_status)) + 
      geom_boxplot() +
      labs(title = paste(drug, "'s Imputed Sensitivity in TCGA BRCA Cohort by TNBC Status", sep = ""),
           y = "Imputed Sensitivity Score") +
      theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5))+
      annotation_custom(grob_annot))
      
}
library(ggpubr)
library(ggplot2)

Sensitivity_grapher <- function(drug, comparison){
print(
  preds.subtype %>% 
    filter(Drug == drug) %>% 
    ggplot(aes(x = eval(parse(text = comparison)), y = PSS, fill = x)) + 
      ggboxplot() +
      labs(title = paste(drug, "'s Imputed Sensitivity in TCGA BRCA Cohort by TNBC Status", sep = ""),
           y = "Imputed Sensitivity Score") +
      theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5))+
    stat_compare_means(method = "anova", label.y = 1)
      
)
  
}


drugs.of.interest <- c("tamoxifen", "fulvestrant", "lapatinib", "MK-1775")

for(i in 1:length(drugs.of.interest)){
  PAM50_ISS_grapher(drugs.of.interest[i])
  Clin.sub_ISS_grapher(drugs.of.interest[i])
  TNBC_ISS_grapher(drugs.of.interest[i])
}


```

```{r}
  grob_annot <- grobTree(textGrob(paste("p-value =", 
                                      format(t.tst_preds.subtype$bonf_p.value[t.tst_preds.subtype$Drug == drug], 
                                             trim = T, digits = 4)), 
                                x = 0.73, y = 0.95, hjust = 0, gp = gpar(fontface = "italic")))
  print(preds.subtype %>% 
    filter(Drug == drug) %>% 
    ggplot(aes(x = TNBC_status, y = PSS, fill = TNBC_status)) + 
      geom_boxplot() +
      labs(title = paste(drug, "'s Imputed Sensitivity in TCGA BRCA Cohort by TNBC Status", sep = ""),
           y = "Imputed Sensitivity Score") +
      theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5))+
      annotation_custom(grob_annot))
```



```{r test function}

library(ggpubr)
library(ggplot2)


Sensitivity_grapher <- function(drug, comparison, group = "TNBC"){
test <- combn(unique(eval(parse(text = paste("preds.subtype$", comparison)))), 2)
test_list <- split(test, rep(1:ncol(test), each = nrow(test)))
comp_list <- test_list[grepl(pattern = group, test_list)]

box_labs <- preds.subtype %>% 
    filter(Drug == drug) %>% 
    count_(comparison) %>% 
    .$n

print(
  preds.subtype %>% 
    filter(Drug == drug) %>% 
    arrange_(comparison) %>% 
  ggboxplot(x = comparison, y = "PSS", fill = comparison) +
      labs(title = paste(drug, "'s Imputed Sensitivity in TCGA BRCA Cohort by TNBC Status", sep = ""),
           y = "Imputed Sensitivity Score", 
           x = "BRCA Clinical Subtype") +
      theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5))+
    stat_compare_means(comparisons = comp_list) + 
    grids() + 
    stat_summary(geom = "text", label = paste("n =", box_labs), vjust = -0.5)
      
)
  
}


Sensitivity_grapher(drug = "MK-1775", comparison = "Clin_Subtype")

Sensitivity_grapher(drug = "lapatinib", comparison = "PAM50", group = "HER2")

```

```{r}
plot_sensitivity2 <- function(.data, .drug = "MK-1775", .grp = "Clin_Subtype", .ref.group = "TNBC"){
  
  #filter
  .data <- filter(.data, Drug == .drug) %>% arrange_(.grp)
  # Compute p-values
  comparison.formula <- paste0("PSS", "~", .grp) %>%
    as.formula()
  pvalues <- compare_means(
    comparison.formula,  data = .data,
    method = "t.test", ref.group = .ref.group
    )
  # P-value y coordinates
  y.max <- .data %>% 
    pull("PSS") %>% max(na.rm = TRUE) 
  p.value.y.coord <- rep(y.max, nrow(pvalues))
  
  step.increase <- (1:nrow(pvalues))*(y.max/40)
  p.value.y.coord <- p.value.y.coord + step.increase
  
  pvalues <- pvalues %>%
    mutate(y.coord =  p.value.y.coord)
  
  box_labs <- preds.subtype %>% 
    filter(Drug == drug) %>% 
    count_(comparison) %>% 
    .$n
  
  ggboxplot(.data, x = .grp, y = "PSS", fill = .grp)+
    ggsignif::geom_signif(
      data = pvalues, 
      aes(xmin = group1, xmax = group2, annotations = p.adj, y_position = y.coord), 
      manual= TRUE, tip_length = 0.03, size = 0.4) + 
    grids() + 
    labs(x = "Clinical Subtype", 
         y = "Predicted Sensitivity Score", 
         title = paste(.drug, "'s Imputed Sensitivity in TCGA BRCA Cohort by TNBC Status", sep = "")) + 
    theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5)) + 
    stat_summary(geom = "text", label = paste("n =", box_labs), vjust = -0.5)
}


# Graph
plot_sensitivity2(.data = preds.subtype, .drug = "MK-1775" , .grp = "Clin_Subtype", .ref.group = "TNBC")

plot_sensitivity2(.data = preds.subtype, .drug = "lapatinib" , .grp = "Clin_Subtype", .ref.group = "HER2+")
```



```{r}
# test <- combn(unique(preds.subtype$Clin_Subtype), 2)
# test_list <- split(test, rep(1:ncol(test), each = nrow(test)))
# tnbc_list <- test_list[grepl(pattern = "TNBC", test_list)]
# 
# 
# drug = "MK-1775"
# comparison = "Clin_Subtype"
# 
# preds.subtype %>% 
#     filter(Drug == drug) %>% 
#     ggboxplot(x = comparison, y = "PSS", fill = comparison) +
#       labs(title = paste(drug, "'s Imputed Sensitivity in TCGA BRCA Cohort by TNBC Status", sep = ""),
#            y = "Imputed Sensitivity Score") +
#       theme(legend.position = "none",
#             plot.title = element_text(hjust = 0.5))+
#       stat_compare_means(comparisons = tnbc_list)
# 
# 
# 
# buld_boxplot <- ggplot_build(boxplot_temp)
# 
# buld_boxplot$data[[2]]$annotation <- c(1:9) #Calculate p-values (TukeyHSD, etc. and put them here)
# 
# plot(ggplot_gtable(buld_boxplot))

```


## Correlations With Molecular Targets

We can look into this further by plotting Tamoxifen and Lapatinib ISS by ER and HER2 expression and seeing how well they correlate. 


```{r ER and HER2 correlations}
###########ER and Tamoxifen Response#################

#ER and Tamoxifen stats:

ER_t.tst_preds <- preds.subtype %>% 
  group_by(Drug) %>% 
  drop_na(er_status_by_ihc) %>% 
  do(tidy(t.test(PSS~er_status_by_ihc, data = .))) %>% 
  ungroup() %>% 
  mutate(bonf_p.value = p.adjust(p.value, method = "bonferroni")) %>% 
  arrange(bonf_p.value) %>% 
  filter(estimate > 0)

head(ER_t.tst_preds)

#ER positivity associates with Imputed Tamoxifen Response, it is the 5th most significant result
```

For ER, we see good correlation between ER status and imputed response to Tamoxifen. Additionally, we see that this imputed repsonse is associated with ER expression.


We can do a similar analysis for HER2. 

```{r}
############### HER2 and Lapatinib Response ##############################


HER2_t.tst_preds <- preds.subtype %>% 
  group_by(Drug) %>% 
  do(tidy(t.test(PSS~HER2_Subtype, data = .))) %>% 
  ungroup() %>% 
  mutate(bonf_p.value = p.adjust(p.value, method = "bonferroni")) %>% 
  arrange(bonf_p.value) %>% 
  filter(estimate > 0)

head(HER2_t.tst_preds)

# HER2 positiivity associates with Lapatinib response, is the 2nd most significant result. 

HER2_PSS <- preds.subtype %>% 
  filter(Drug == "lapatinib") %>% 
  drop_na(ERBB2_CN) %$%
  cor.test(PSS, ERBB2_CN, method = "spearman")

preds.subtype %>% 
  filter(Drug == "lapatinib") %>% 
  ggplot(aes(x = ERBB2_CN, y = PSS)) + 
    geom_point() +
    geom_smooth(method = "lm") + 
    annotate("text", x = 2, y = 14.1, label = paste("p-value =", HER2_PSS$p.value, ",  rho =", HER2_PSS$estimate, sep = "")) + 
    labs(title = "Imputed Lapatinib Response Correlates \n with HER2 Copy Number in TCGA BRCA Patients", 
         x = "ERBB2 Copy Number", 
         y = "Lapatinib Imputed Sensitivity Score (ISS)") + 
    theme(plot.title = element_text(hjust = 0.5))

```


For HER2, we see good correlation between ER status and imputed response to Tamoxifen. Additionally, we see that this imputed repsonse is associated with ER expression.



## TNBC Analysis 

We want to pick drugs that are better in TNBC by both effect size and p-value cutoffs. For this, we will use the top 20 most significant hits that have an effect size at least as great as taxomifen in ER+ cancers. Here is the list of drugs (with one drug combo) that fits this criteria


```{r AOV Beta analysis}
#perform the effect size analysis to get eta-squared values
b_preds.subtype <- preds.subtype %>% 
  group_by(Drug) %>% 
  do(tidy(lsr::etaSquared(aov(.$PSS ~ .$PAM50))))

#we will also want to perform a Tukey HSD so that we can determine which group is the one that is most effective
HSD_preds.subtype <- preds.subtype %>% 
  group_by(Drug) %>% 
  do(tidy(TukeyHSD(aov(.$PSS ~ .$PAM50)))) 

#now joining the two datasets and for each drug determining which is the sensitive group
eta.sq_preds.subtype <- filter(b_preds.subtype) %>% 
  left_join(x= . , y = HSD_preds.subtype) %>% 
  group_by(Drug) %>% 
  summarise(sensitive_group = if_else(sum(estimate[1:4] > 0) == 4, 
                                      true = "basal", 
                                      false = if_else(sum(c(estimate[1] < 0,estimate[5:7] > 0)) == 4, 
                                                      true = "HER2", 
                                                      false = if_else(sum(estimate[c(2:3, 5:6)] < 0) == 4, 
                                                                      true = "luminal", 
                                                                      false = "none"))),
            eta.sq = eta.sq[1]) %>% 
  left_join(., filter(aov_preds.subtype, term == ".$PAM50"))

eta.sq_preds.subtype %>% 
  filter(eta.sq > eta.sq[Drug == "tamoxifen"], 
         sensitive_group == "basal")


```


Now that we have a drug list, we can identify what pathways these drugs target. The list indicates that cell cycle is the most prominent pathway targeted. 

SAVE THE DRUG FILE TO USE IN ANALYZE IDWAS RESULTS!!!


```{r drug targets}

drug_targets <- read_csv("Drug_targets(review).csv") # From Ling and Gruener et al

targets_eta.sq <- eta.sq_preds.subtype %>% 
  filter(eta.sq > eta.sq[Drug == "tamoxifen"], 
         sensitive_group == "basal") %>% 
  select(Drug, sensitive_group, eta.sq, bonf_p.value) %>% 
  left_join(., drug_targets, by = c("Drug" = "Name in Database")) %>% 
  select(-c(7:9)) %>% 
  distinct() 

#Using the review information as well some manual curation, these are the simplified pathways for these compounds 
simplified_paths <- c("Amino Acid Metabolism", "Cell Cycle", "Cell Cycle", "Protein Stability and Degradation", "Cell Cycle", "Unknown", "Cell Cycle", "Sphingolipid Metabolism", "Cell Cycle", "Cell Cycle", "Protein Stability and Degradation", "STAT Signaling", "Cell Cycle", "Notch Signaling", "Purine Metabolism")

names(simplified_paths) <- c("968", "AZD7762", "BI-2536", "bortezomib", "BRD-K30748066", "BRD-K66453893", "BRD9876:MK-1775 (4:1 mol/mol)", "ceranib-2", "GSK461364", "MK-1775", "MLN2238", "NSC 74859", "PHA-793887", "semagacestat", "triazolothiadiazine")

targets_eta.sq <- targets_eta.sq %>% 
  mutate(Pathway = simplified_paths[names(simplified_paths) == Drug]) %>% 
  select(1:4, Pathway)

targets_eta.sq

```

We can go further and use the entire TCGA patient data and find associations between genetic features (i.e. non-synonymous mutations) and the patient's Imputed Sensitivty Scores. If we graph the 9 genes with the most significant values, this is the result.  



```{r}
t.tst_preds.subtype %>% mutate(Sensitive_Group = if_else(estimate > 0, "TNBC", "RPBC")) %>% 
  select(Drug, Sensitive_Group, bonf_p.value) %>%  arrange(bonf_p.value) 
```

