---
title: "Analyze IDWAS Results"
output: html_notebook
---

Now that we have a drug list, we can identify what pathways these drugs target. The list indicates that cell cycle is the most prominent pathway targeted. 

#need to load data from Analyze_TCGA-Imputation-Data as well as the currently non-existant Impute_All_TCGA_for_IDWAS.Rmd and IDWAS_Workflow.Rmd


We can go further and use the entire TCGA patient data and find associations between genetic features (i.e. non-synonymous mutations) and the patient's Imputed Sensitivty Scores. If we graph the 9 genes with the most significant values, this is the result.  


```{r Biomarker Analysis}

library(tidyverse)
theme_set(theme_bw())
TCGA_BRCA_muts_idwa <- read_csv("E:/Research/a) General-methods_Imputations_Biomarkers/b) IDWAS and Biomarkers/IDWAS/tables/CTRPResults_BRCAonly.csv")

IDWAS_results <-  TCGA_BRCA_muts_idwa %>% 
  separate(col = X1, into = c("Drug", "Gene"), sep = "\\.") %>% 
  rename(FDR = `FDR (BH corrected)`) %>% 
  mutate(logFDR = -log10(FDR))

IDWAS_results
```

```{r Drug Based Analysis}
azd_p53<- IDWAS_results %>% 
  filter(Drug == "MK-1775") %>% 
  ggplot(aes(x = logFDR)) + 
    geom_histogram(bins = 20, fill = "light blue", color = "black") +
    labs(y = "Number of Associations", 
         x = "Significance (-log10 FDR)") +
    xlim(NA, 48) +
  geom_vline(xintercept = 45.9, color = "red") +
  annotate(geom = "text", x = 34, y = 200, parse = T, 
           label = as.character(expression(atop("p53-AZD1775 Association", "(red line, FDR = 1.2x10"^{-46}*")"))))

azd_p53

IDWAS_results %>% 
  filter(Drug == "MK-1775", 
         FDR < 0.01)

IDWAS_results %>% 
  filter(FDR < 0.0001) %>% 
  count(Gene) %>% arrange(desc(n))

```


```{r Gene-Based Analysis}

TCGA_muts_idwa <- read_csv("E:/Research/a) General-methods_Imputations_Biomarkers/b) IDWAS and Biomarkers/IDWAS/tables/CTRP_TCGAmuts_allResults_controlFORtissue.csv")


table_top_drugs <- read_csv("./DataOut/Table1_draft.csv")

top_drugs <- table_top_drugs %>% 
  separate_rows(drugs, sep = ", ") %>% 
  rename(Drug = drugs)

genes_top <- TCGA_muts_idwa %>% 
  separate(col = X1, into = c("Drug", "Gene"), sep = "\\.") %>%
  rename(FDR = `FDR (BH corrected)`) %>% 
  mutate(logFDR = -log10(FDR)) %>% 
  mutate(Drug = replace(Drug, Drug == "MK-1775", "AZD-1775")) %>% 
  right_join(., top_drugs, by = "Drug")

write_csv(genes_top, path = "./DataOut/Supplemental_table5.csv")

graph_FDR_cutoff <- genes_top %>% 
  drop_na(Gene) %>% 
  filter(.$Gene %in% unique(.$Gene[order(.$FDR)])[1:12]) %>% #written this way, gets the top 12 (arbitrary, just for aesthetics) unique genes based on FDR value, then filters the dataset to keep these genes for all the drugs
  mutate(Gene = factor(x = Gene, levels = unique(.$Gene[order(.$FDR)])[1:12]))

graph_FDR_cutoff %>% 
  mutate(color = if_else(Drug == "AZD-1775", "AZD-1775",
                         if_else(MOA %in% c("CDK inhibitor", "CHK inhibitor", "PLK inhibitor"), "Cell Cycle", "Other"))) %>% 
  ggplot(mapping = aes(x = `Effect Size (beta)`, y = logFDR, color = color)) + 
    geom_point() + 
    facet_wrap(~Gene, scales = "free_y") +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") + 
    geom_hline(yintercept = -log10(0.01), color = "red") + 
    labs(title = "Top Mutation-Imputed-Drug Associations in TCGA",
         y = "-log(FDR)") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_color_manual(values = c("red", "light blue2", "black"))

sup3a <- graph_FDR_cutoff %>% 
  mutate(`Mechanism of Action` = if_else(Drug == "AZD-1775", " Wee1 Inhibition",
                         if_else(MOA %in% c("CDK inhibitor", "CHK inhibitor", "PLK inhibitor"), "Cell Cycle", "Other"))) %>% 
  ggplot(mapping = aes(x = `Effect Size (beta)`, y = logFDR, color = `Mechanism of Action`)) + 
    geom_point() + 
    facet_wrap(~Gene, scales = "free_y") +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") + 
    geom_hline(yintercept = -log10(0.01), color = "red") + 
    labs(y = "-log(FDR)") + 
    scale_color_manual(values = c("red", "#2471A3", "black")) + 
    theme(legend.position = "top")


```


```{r}
graph_FDR_cutoff %>% filter(logFDR > 5) %>% arrange(FDR)

```




##### Supplmental Code

```{r Gene-based only BRCA}

brca_genes_top <- IDWAS_results %>% 
  mutate(Drug = replace(Drug, Drug == "MK-1775", "AZD-1775")) %>% 
  right_join(., top_drugs, by = "Drug")

brca_graph_FDR_cutoff <- brca_genes_top %>% 
  drop_na(Gene) %>% 
  filter(.$Gene %in% unique(.$Gene[order(.$FDR)])[1:12]) %>%  #written this way, gets the top 12 (arbitrary, just for aesthetics) unique genes based on FDR value, then filters the dataset to keep these genes for all the drugs
  mutate(Gene = factor(x = Gene, levels = unique(.$Gene[order(.$FDR)])[1:12]))

brca_graph_FDR_cutoff %>% 
  mutate(color = if_else(Drug == "AZD-1775", "AZD-1775",
                         if_else(MOA %in% c("CDK inhibitor", "CHK inhibitor", "PLK inhibitor"), "Cell Cycle", "Other"))) %>% 
  ggplot(mapping = aes(x = `Effect Size (beta)`, y = logFDR, color = color)) + 
    geom_point() + 
    facet_wrap(~Gene, scales = "free_y") +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") + 
    geom_hline(yintercept = -log10(0.01), color = "red") + 
    labs(title = "Top Mutation-Imputed-Drug Associations in TCGA BRCA",
         y = "-log(FDR)") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_color_manual(values = c("red", "light blue2", "black"))

```




```{r}
all_tcga_azd1775 <- TCGA_muts_idwa %>% 
  separate(col = X1, into = c("Drug", "Gene"), sep = "\\.") %>%
  rename(FDR = `FDR (BH corrected)`) %>% 
  mutate(logFDR = -log10(FDR)) %>% 
  filter(Drug == "MK-1775")

all_tcga_azd1775

all_tcga_azd1775 %>% 
  ggplot(aes(x = logFDR)) + 
    geom_histogram(bins = 20, fill = "light blue", color = "black") +
    labs(y = "Number of Associations", 
         x = "Significance (-log10 FDR)")

```


