---
title: "Analyze IDWAS Results"
output: html_notebook
author:
  - "Robert F. Gruener"
date: "date last edited: `r format(Sys.time(), '%d %B, %Y')`"
---
```{r, include=FALSE}
knitr::opts_chunk$set(warning = F, message = FALSE, cache = T)
```


Now that we have a drug list, we can identify what pathways these drugs target. The list indicates that cell cycle is the most prominent pathway targeted. 

#need to load data from Analyze_TCGA-Imputation-Data as well as the currently non-existant Impute_All_TCGA_for_IDWAS.Rmd and IDWAS_Workflow.Rmd


We can go further and use the entire TCGA patient data and find associations between genetic features (i.e. non-synonymous mutations) and the patient's Imputed Sensitivty Scores. If we graph the 9 genes with the most significant values, this is the result.  


```{r Biomarker Analysis}

library(tidyverse)
theme_set(theme_bw())
TCGA_BRCA_muts_idwa <- read_csv("./Included_data/CTRPResults_BRCAonly.csv")

IDWAS_results <-  TCGA_BRCA_muts_idwa %>% 
  separate(col = X1, into = c("Drug", "Gene"), sep = "\\.") %>% 
  rename(FDR = `FDR (BH corrected)`) %>% 
  mutate(logFDR = -log10(FDR))

IDWAS_results
```

```{r Drug Based Analysis}
azd_p53<- IDWAS_results %>% 
  filter(Drug == "MK-1775") %>% 
  ggplot(aes(x = logFDR)) + 
    geom_histogram(bins = 20, fill = "light blue", color = "black") +
    labs(y = "Number of Associations", 
         x = "Significance (-log10 FDR)") +
    xlim(NA, 48) +
  geom_vline(xintercept = 45.9, color = "red") +
  annotate(geom = "text", x = 34, y = 200, parse = T, 
           label = as.character(expression(atop("p53-AZD1775 Association", "(red line, FDR = 1.2x10"^{-46}*")"))))

azd_p53

IDWAS_results %>% 
  filter(Drug == "MK-1775", 
         FDR < 0.01)

IDWAS_results %>% 
  filter(FDR < 0.0001) %>% 
  count(Gene) %>% arrange(desc(n))

```

```{r}
ggsave(plot = azd_p53, filename = "./Figures/Figure3b.png")
```


```{r Top 12 Gene-Based Analysis}

TCGA_muts_idwa <- IDWAS_results

table_top_drugs <- read_csv("./DataOut/Table1_draft.csv")

top_drugs <- table_top_drugs %>% 
  separate_rows(drugs, sep = ", ") %>% 
  rename(Drug = drugs)

genes_top <- TCGA_muts_idwa %>% 
  mutate(Drug = replace(Drug, Drug == "MK-1775", "AZD-1775")) %>% 
  right_join(., top_drugs, by = "Drug") %>% 
  drop_na(Gene) %>% 
  mutate(color = if_else(Drug == "AZD-1775", "AZD-1775",
                         if_else(MOA %in% c("CDK inhibitor", "CHK inhibitor", "PLK inhibitor"), "Cell Cycle", "Other")))

write_csv(genes_top, path = "./DataOut/Supplemental_table5.csv")

graph_FDR_cutoff <- genes_top %>% 
  drop_na(Gene) %>% 
  filter(Gene %in% unique(.$Gene[order(.$FDR)])[1:12]) %>% #written this way, gets the top 12 (arbitrary, just for aesthetics) unique genes based on FDR value, then filters the dataset to keep these genes for all the drugs
  mutate(Gene = factor(x = Gene, levels = unique(.$Gene[order(.$FDR)])[1:12])) %>% 
  mutate(color = if_else(Drug == "AZD-1775", "AZD-1775",
                         if_else(MOA %in% c("CDK inhibitor", "CHK inhibitor", "PLK inhibitor"), "Cell Cycle", "Other")))

sup3a <- graph_FDR_cutoff  %>% 
  ggplot(mapping = aes(x = `Effect Size (beta)`, y = logFDR, color = color)) + 
    geom_point() + 
    facet_wrap(~Gene, scales = "free_y") +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") + 
    geom_hline(yintercept = -log10(0.05), color = "red") + 
    labs(y = "-log(FDR)") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_color_manual(values = c("red", "#2471A3", "black")) + 
    theme(legend.position = "top")

sup3a 


```


```{r All Significant Gene-Based Analysis}

graph_FDR_cutoff <- genes_top %>% 
  filter(Gene %in% unique(.$Gene[.$FDR < 0.05])) %>% 
  mutate(Gene = factor(x = Gene, levels = unique(.$Gene[order(.$FDR)]))) 

test_graph_allsigs <- graph_FDR_cutoff  %>% 
  ggplot(mapping = aes(x = `Effect Size (beta)`, y = logFDR, color = color)) + 
    geom_point() + 
    facet_wrap(~Gene, scales = "free_y") +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") + 
    geom_hline(yintercept = -log10(0.05), color = "red") + 
    labs(y = "-log(FDR)") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_color_manual(values = c("red", "#2471A3", "black")) + 
    theme(legend.position = "top")

test_graph_allsigs 


```


```{r}
genes_top %>% filter(logFDR > -log10(0.05)) %>% arrange(Drug) %>% 
  group_by(Gene) %>% mutate(Drugs = paste(Drug, collapse = ", ")) %>% count(Gene, Drugs) %>% select(Gene, n, Drugs) %>% knitr::kable(.)

genes_top %>% filter(logFDR > -log10(0.05)) %>% arrange(Gene) %>% 
  group_by(Drug) %>% mutate(Genes = paste(Gene, collapse = ", ")) %>% count(Genes, Drug) %>%  select(Drug, n, Genes) %>% 
  knitr::kable(.)

genes_top %>% filter(logFDR > -log10(0.05)) %>% arrange(Gene) %>% 
  group_by(Drug) %>% mutate(Genes = paste(Gene, collapse = ", ")) %>% count(Genes, Drug, MOA) %>%  select(Drug, n, Genes, MOA) %>% 
  arrange(MOA) %>%  knitr::kable(.)

genes_top %>% filter(logFDR > -log10(0.05), color %in% c("AZD-1775", "Cell Cycle")) %>% arrange(Drug) %>% 
  group_by(Gene) %>% mutate(Drugs = paste(Drug, collapse = ", ")) %>% count(Gene, Drugs) %>%  select(Gene, n, Drugs) %>% 
  arrange(desc(n)) %>%  knitr::kable(.)

```


```{r}
genes_top %>% filter(logFDR > -log10(0.05)) %>% arrange(Gene) %>% mutate(drug_spread = Drug) %>% select(drug_spread, Drug, Gene) %>% 
   spread(key = Gene, value = drug_spread) 

genes_top %>% filter(logFDR > -log10(0.05)) %>% arrange(Gene) %>% select(MOA, Drug, Gene) %>% 
   spread(key = Gene, value = MOA)
  

genes_top %>% filter(logFDR > -log10(0.05)) %>% arrange(Gene) %>% select(MOA, Drug, Gene, color) %>% 
  ggplot(aes(x = Drug, y = Gene, fill = color)) +
  geom_tile() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



```






Only include the genes that significantly associate with AZD-1775
```{r AZD-1775 Focused Figure}
graph_FDR_cutoff <- genes_top %>% 
  filter(FDR < 0.047, Drug == "AZD-1775") %>% select(Gene) %>% left_join(.,genes_top) %>% 
  mutate(Gene = factor(x = Gene, levels = unique(.$Gene[order(.$FDR)]))) 

azd1775_graph_allsigs <- graph_FDR_cutoff  %>% 
  ggplot(mapping = aes(x = `Effect Size (beta)`, y = logFDR, color = color)) + 
    geom_point() + 
    facet_wrap(~Gene, scales = "free_y") +
    scale_y_continuous(limits=c(0,NA)) +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") + 
    geom_hline(yintercept = -log10(0.05), color = "red") + 
    labs(y = "-log(FDR)") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_color_manual(values = c("red", "#2471A3", "black")) + 
    theme(legend.position = "top")

azd1775_graph_allsigs 
```

```{r}
ggsave(plot = azd1775_graph_allsigs, filename = "./Figures/Supplemental_Figure_3.png", height = 5.61, width = 8.68)
```



```{r Only cell cycle MOA}
genes_top %>% filter(Gene %in% unique(.$Gene[.$FDR < 0.05]), color %in% c("AZD-1775", "Cell Cycle")) %>%   
  mutate(Gene = factor(x = Gene, levels = unique(.$Gene[order(.$FDR)]))) %>% 
  ggplot(mapping = aes(x = `Effect Size (beta)`, y = logFDR, color = MOA)) + 
    geom_point() + 
    facet_wrap(~Gene, scales = "free_y") +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") + 
    geom_hline(yintercept = -log10(0.05), color = "red") + 
    labs(y = "-log(FDR)") + 
    theme(plot.title = element_text(hjust = 0.5)) +  
    theme(legend.position = "top")
  
  
```


##### Supplmental Code

```{r Gene-based only BRCA}

brca_genes_top <- IDWAS_results %>% 
  mutate(Drug = replace(Drug, Drug == "MK-1775", "AZD-1775")) %>% 
  right_join(., top_drugs, by = "Drug")

brca_graph_FDR_cutoff <- brca_genes_top %>% 
  drop_na(Gene) %>% 
  filter(.$Gene %in% unique(.$Gene[order(.$FDR)])[1:12]) %>%  #written this way, gets the top 12 (arbitrary, just for aesthetics) unique genes based on FDR value, then filters the dataset to keep these genes for all the drugs
  mutate(Gene = factor(x = Gene, levels = unique(.$Gene[order(.$FDR)])[1:12]))

brca_graph_FDR_cutoff %>% 
  mutate(color = if_else(Drug == "AZD-1775", "AZD-1775",
                         if_else(MOA %in% c("CDK inhibitor", "CHK inhibitor", "PLK inhibitor"), "Cell Cycle", "Other"))) %>% 
  ggplot(mapping = aes(x = `Effect Size (beta)`, y = logFDR, color = color)) + 
    geom_point() + 
    facet_wrap(~Gene, scales = "free_y") +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") + 
    geom_hline(yintercept = -log10(0.01), color = "red") + 
    labs(title = "Top Mutation-Imputed-Drug Associations in TCGA BRCA",
         y = "-log(FDR)") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_color_manual(values = c("red", "light blue2", "black"))

```




```{r}
all_tcga_azd1775 <- TCGA_muts_idwa %>% 
  separate(col = X1, into = c("Drug", "Gene"), sep = "\\.") %>%
  rename(FDR = `FDR (BH corrected)`) %>% 
  mutate(logFDR = -log10(FDR)) %>% 
  filter(Drug == "MK-1775")

all_tcga_azd1775

all_tcga_azd1775 %>% 
  ggplot(aes(x = logFDR)) + 
    geom_histogram(bins = 20, fill = "light blue", color = "black") +
    labs(y = "Number of Associations", 
         x = "Significance (-log10 FDR)")

```

```{r}
print(sessionInfo())
```
