---
title: "Download GDSC and CTRP Data"
output: html_notebook
---

## Objectives:

The goal of this file is to: download and install the necessary packages and files into the R project. Specifically: 

 * Install pRRophetic and it's dependencies
 * Download the cell line screening data
 * Download the RNA-Seq information
 * Download the TCGA dataset
 
Important note, certain files are already included in this parent directory including "pRRophetic_0.5.tar.gz" , "All Cell Line Infor.xlsx" , and  "Breast_CCL_Subtypes.csv". Particularly for the first two files listed, they are needed to run the imputations, the last for Analyzing the imputation data in GDSC. 

pRRophetic_0.5.tar.gz can be downloaded here
All Cell Line Infor.xlsx is already included in this github repo, it was originally published in our review doi:
Breast_CCL_Subtypes.csv is already included in this github repo, it was originally part of this publication citation, doi:, supplemental table X

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```


### Installing required packages (pRRophetic must be downloaded and put in wd or project home)

Again to install pRRophetic, it will need to be downloaded and placed in the repo manually/separately. 

```{r install and load pRRophetic and dependencies, results= "hide"}

#install BiocManager if needed
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
#install pRRophetic dependencies if needed
if (!requireNamespace(c("car", "ridge", "preprocessCore", "genefilter", "sva"), quietly = TRUE)) 
    BiocManager::install(c("car", "ridge", "preprocessCore", "genefilter", "sva"))

#install pRRophetic from tar.gz file (place in repo seperately)
if (!requireNamespace("pRRophetic", quietly = TRUE)) 
    install.packages("pRRophetic_0.5.tar.gz", repos = NULL, type = "source")

#install other packages that will be used in this script
if (!requireNamespace(c("qvalue", "affy", "GenomicFeatures", "org.Hs.eg.db", "biomaRt" , 
                        "TxDb.Hsapiens.UCSC.hg19.knownGene", "glmnet", "gdata", "tidyverse", 
                        "illuminaHumanv4.db", "broom", "ggpubr"), quietly = TRUE)) 
    BiocManager::install(c("qvalue", "affy", "GenomicFeatures", "org.Hs.eg.db", "biomaRt" , 
                        "TxDb.Hsapiens.UCSC.hg19.knownGene", "glmnet", "gdata", "tidyverse", 
                        "illuminaHumanv4.db", "broom", "ggpubr"))

dir.create("./DataIn")

```

### Downloading Cell Line Data


#### Download the data from GDSC. 

NB: Can find the GDSC expression data [at this link](https://www.cancerrxgene.org/gdsc1000/GDSC1000_WebResources//Data/preprocessed/Cell_line_RMA_proc_basalExp.txt.zip )
However, this information wasn't updated any since 2016 and the 2016 data is already packaged into this version of pRRophetic as cgp2016ExprRma, so we will just call the data from pRRophetic. 

```{r downloading GDSC IC50 data}
dir.create("./DataIn/GDSC")
download.file(url = "ftp://ftp.sanger.ac.uk/pub/project/cancerrxgene/releases/current_release/GDSC2_fitted_dose_response_17Jul19.xlsx",
                destfile = "./DataIn/GDSC/GDSC_IC50_July2019.xlsx", mode = "wb")

```


Downloads the data from CTRP and CCLE. Data can also be accessed from these links:

CTRPv2 Data from: ftp://caftpd.nci.nih.gov/pub/OCG-DCC/CTD2/Broad/CTRPv2.0_2015_ctd2_ExpandedDataset/
CCLE Data from: https://portals.broadinstitute.org/ccle/data


```{r downloading CTRP and CCLE data}
dir.create("./DataIn/CTRPv2")
dir.create("./DataIn/CCLE")

#CTRP Data
temp.zip <- tempfile()
download.file(url = "ftp://caftpd.nci.nih.gov/pub/OCG-DCC/CTD2/Broad/CTRPv2.0_2015_ctd2_ExpandedDataset/CTRPv2.0_2015_ctd2_ExpandedDataset.zip",
              "./DataIn/CTRP.zip") 
unzip("./DataIn/CTRP.zip", exdir = "./DataIn/CTRPv2")
unlink("./DataIn/CTRP.zip")

#Download CCLE Data 
download.file(url = "https://data.broadinstitute.org/ccle/CCLE_DepMap_18Q2_RNAseq_RPKM_20180502.gct", 
              destfile = "./DataIn/CCLE/CCLE_RPKM.gct")

######## Commenting Out other Download Files That Alex Used, since these files will not be used

# download.file(url = "https://data.broadinstitute.org/ccle/CCLE_DepMap_18Q2_RNAseq_reads_20180502.gct", 
#        destfile = "./DataIn/CCLE/CCLE_RNAseq_reads.gct")
# download.file(url = "https://data.broadinstitute.org/ccle_legacy_data/mRNA_expression/CCLE_Expression_Entrez_2012-09-29.gct", 
#              destfile = "./DataIn/CCLE/CCLE_ExpressionEntrez.gct")


```

### Downloading the TCGA data from FireBrowse


```{r downloading TCGA data}
#this should be updated to the GDC instead of Firebrowse at some point
dir.create("./DataIn/TCGA")
download.file("http://gdac.broadinstitute.org/runs/stddata__2015_08_21/data/BRCA/20150821/gdac.broadinstitute.org_BRCA.Merge_rnaseqv2__illuminahiseq_rnaseqv2__unc_edu__Level_3__RSEM_genes__data.Level_3.2015082100.0.0.tar.gz", "./DataIn/TCGA/TCGA_BRCA_RNA-seq.tar.gz")
untar("./DataIn/TCGA/TCGA_BRCA_RNA-seq.tar.gz", exdir = "./DataIn/TCGA/")
#having issues with the untar file name being too long, renaming
file.rename(from = "./DataIn/TCGA/gdac.broadinstitute.org_BRCA.Merge_rnaseqv2__illuminahiseq_rnaseqv2__unc_edu__Level_3__RSEM_genes__data.Level_3.2015082100.0.0/", to = "./DataIn/TCGA/RNA-seq" )
unlink("./DataIn/TCGA/TCGA_BRCA_RNA-seq.tar.gz")


```
