---
title: 'Imputations: Build models using the CCLE/CTRPv2 datasets to impute drug response in TCGA BRCA Tumors'
output: html_notebook 
---

### Objectives

Do the imputations

 * Set the parameters
 * Load the test expression data (TCGA Breast Cancer)
 * Load the CTRP/CCLE Training expression and drug data
 * Generate the models and impute using the pRRophetic framework 

Important: Set the parameters appropriately, then everything else *should* run smoothly if you run the enitre file.  

### First, load packages and set parameters
```{r loading packages and setting parameters, results= "hide"}
set.seed(12345)
#Loading Packages
needed_packages <- c("tidyverse", "qvalue", "affy", "GenomicFeatures", "org.Hs.eg.db", "TxDb.Hsapiens.UCSC.hg19.knownGene", 
                     "parallel", "biomaRt", "readxl", 
                     "glmnet", "gdata", "illuminaHumanv4.db" ,"car", "ridge", "preprocessCore", "genefilter", "sva", "pRRophetic")
lapply(needed_packages, require, character.only = TRUE)

###################### Set Parameters ##########################
# Set location and filename  for list/matrix R.data output (ctrp_list_output_name, ctrp_mat_output_name)

dir.create("DataOut")
matrix_output_name <- "./DataOut/CTRP-Models_on_TCGA-BRCA-RNA.RData"

# Finally, set parameters of pRRophetic's calcphenotype function 
#batchCorrect options: "eb" for ComBat, "qn" for quantiles normalization, "standardize", or "none"
batchCorrect <- "eb"
#powerTransformPhenotype: does a power transformation to the phenotype data, recommended over log transformations by the literture 
powerTransformPhenotype <- TRUE
#removeLowVaryingGenes: percentage of low varying genes to remove, default is at 0.2 
removeLowVaryingGenes <- 0.2
#removeLowVaryingGenesFrom: where should it remove low varying genes from "homogenizeData" or "rawData"
removeLowVaringGenesFrom <- "homogenizeData"
#minNumSamples: minimum number of training samples, set to 30, but this shouldn't be an issue since the there are many samples 
minNumSamples = 30
#selection: how to deal with duplicated gene IDs, -1 for ask user, 1 for summarize by mean, 2 for disregard duplicates
selection <- 1
#printoutput: should it print outputs, default is TRUE
printOutput = TRUE


```

### Second, load the RNA-Seq data

```{r Load RNA-Seq}
############ Load Test Expression data (as Matrix) ###################

# load  the breast cancer RNA-seq file downloaded previously
tpmDatMat <- read.delim(file = "./DataIn/TCGA/RNA-seq/BRCA.rnaseqv2__illuminahiseq_rnaseqv2__unc_edu__Level_3__RSEM_genes__data.data.txt", as.is=T)
tpmDatMat_tpm <- tpmDatMat[-1,which(tpmDatMat[1,] == "scaled_estimate")]
tpmDatMat_tpm <- apply(tpmDatMat_tpm, 2, as.numeric)
geneNames <- do.call(cbind, strsplit(tpmDatMat[, "Hybridization.REF"], "|", fixed=TRUE))[1,][-1]
rownames(tpmDatMat_tpm) <- geneNames
colnames(tpmDatMat_tpm) <- substr(colnames(tpmDatMat_tpm), 1, 28)
tpmDatMat_tpm_logged <- log((tpmDatMat_tpm*1000000)+1) # transform the data
#remove adjacent normal RNA-seq samples
testExprData <- tpmDatMat_tpm_logged[,-grep(pattern = ".11A.|.11B.|.10A.|.10B.", colnames(tpmDatMat_tpm_logged))]

```

### Third, perform Imputations:

Perform imputations for GDSC or CTRP

```{r CTRP Imputations, message = FALSE}

#Loads training phenotype data 
CTRPv2_AUC <- read.delim("./DataIn/CTRPv2/CTRPv2_AUC_clean.txt", sep = "\t", stringsAsFactors = F)
#Altering cell line names that start with a number to start with an X to match rownames of expression data
start.with.number <- grep("^[[:digit:]]", CTRPv2_AUC$CCL_Name)
CTRPv2_AUC$CCL_Name[start.with.number] <- paste("X", CTRPv2_AUC$CCL_Name[start.with.number], sep = "")
#making training PhenData with same layout between CTRP/GDSC ##NB not the same layout
trainingPhenData <- CTRPv2_AUC
#names(trainingPhenData)[names(trainingPhenData)== "cpd_name"] <- "Drug.name"
possibleDrugs <- unique(trainingPhenData$cpd_name)

#Load Training Expression data, convert it to a matrix
CTRPv2_RNAseq_TPM <- read.delim("./DataIn/CTRPv2/CTRPv2_RNAseq_TPM_clean.txt",  sep = "\t", stringsAsFactors = F)
convert.to.matrix <- function(df){
  mat <- as.matrix(df[,-1])
  rownames(mat) <- c(t(df[,1]))
  return(mat)
  }
ExprData <- convert.to.matrix(CTRPv2_RNAseq_TPM)
#make a list of all possible drugs predictions
CTRPDrugPredictions <- list()

########make a for loop for each drug seperate#################
for(i in 1:length(possibleDrugs)) { 
  drug <- possibleDrugs[i]
  temp <- trainingPhenData[trainingPhenData$cpd_name == drug,] ## Makes temp matrix to extract the AUCs easier
  AUCs <- as.numeric(temp$Avg_AUC)
  names(AUCs) <- temp$CCL_Name
  commonCellLines <- colnames(ExprData)[colnames(ExprData) %in% names(AUCs)]
  AUCsOrd <- AUCs[commonCellLines]
  trainDataOrd <- ExprData[, commonCellLines]
  trainDataOrd <- trainDataOrd[(rowSums(trainDataOrd) != 0), ]

  CTRPDrugPredictions[[1]] <- calcPhenotype(trainingExprData =  trainDataOrd, 
                                trainingPtype = AUCsOrd, testExprData = testExprData, batchCorrect = batchCorrect, 
                                powerTransformPhenotype = powerTransformPhenotype, removeLowVaryingGenes = removeLowVaryingGenes, 
                                minNumSamples = minNumSamples, selection = selection, 
                                printOutput = printOutput, removeLowVaringGenesFrom = removeLowVaringGenesFrom)
} #close for loop

names(CTRPDrugPredictions) <- possibleDrugs
CTRPDrugPredictions_mat <- do.call(rbind, CTRPDrugPredictions)
save(CTRPDrugPredictions_mat, file= matrix_output_name)


```



